Coding theory package for {\it Macaulay2}

### 2021-05-13 tex/paper.tex
Fixed issue caused by differences between Macaulay2 1.16 and 1.17, fixed a few minor formatting problems, removed tracking from .dvi and .pdf files
May 13

\begin{document}
\title{Coding theory package for Macaulay2}
%%%%%%%%%%%%%%%%%%%%%%%
\author{Taylor Ball}
\address[Taylor Ball]{Department of Mathematics\\University of Notre Dame\\ Notre Dame, IN USA}
\email{trball13@gmail.com}
%%%%%%%%%%%%%%%%%%%%%%%
\author{Eduardo Camps}
\address[Eduardo Camps]{Escuela Superior de F\'isica y Matem\'aticas \\ Instituto Polit\'ecnico Nacional\\ Mexico City, Mexico}
\email{camps@esfm.ipn.mx}
%%%%%%%%%%%%%%%%%%%%%%%
\author{Henry Chimal-Dzul}
\address[Henry Chimal-Dzul]{Department of Mathematics and Center of Ring Theory and its Applications\\ Ohio University\\ Athens, OH USA}
\email{hc118813@ohio.edu}
%%%%%%%%%%%%%%%%%%%%%%%
\author{Delio Jaramillo-Velez}
\address[Delio Jaramillo-Velez]{
Departamento de
Matem\'aticas\\
Centro de Investigaci\'on y de Estudios
Avanzados del
IPN\\
Apartado Postal
14--740 \\
07000 Mexico City, D.F.
}

@@

\begin{abstract}
In this Macaulay2 \cite{M2} package we define an object called {\it linear code}. We implement functions that compute basic parameters and objects associated with a linear code, such as generator and parity check matrices, the dual code, length, dimension, and minimum distance, among others. We define an object {\it evaluation code}, a construction which allows to study linear codes using tools of
algebraic geometry and commutative algebra. We implement functions to generate important families of linear codes such as Hamming codes, cyclic codes, Reed--Solomon codes, Reed--Muller codes, Cartesian codes, monomial--Cartesian codes, and toric codes. In addition, we define functions for the syndrome decoding algorithm and locally recoverable code construction, which are important tools in applications of linear codes. The package \textit{CodingTheory.m2} is available at \url{https://github.com/Macaulay2/Workshop-2020-Cleveland/tree/CodingTheory/CodingTheory}
\end{abstract}

@@

In this note we introduce a new package written for Macaulay2 \cite{M2} called \textit{CodingTheory}. The goal of this package is to provide a range of functions for constructing  linear and evaluation codes over finite fields, and for computing some of their main properties. To this aim, we define two objects, namely \textit{linear code} and \textit{evaluation code}. The package also includes implementation of functions for generating important families of linear codes like Hamming codes, cyclic codes, Reed-Solomon codes, Reed-Muller codes, Cartesian codes, monomial-Cartesian codes and toric codes. It also has functions for the syndrome decoding algorithm and locally recoverable codes.

The organization of this note is as follows. In Section~\ref{def} we describe different ways to define a linear code over a finite field using the \textit{CodingTheory} package. In Section~\ref{pro} we show how to compute the main parameters of a linear code: length, dimension, and minimum distance. We also illustrate how to compute some of the main algebraic objects associated with linear codes like generator and parity check matrices, dual codes, etc. In Section~\ref{eva} we give a brief introduction to evaluation codes and describe some functions implemented to study these objects. In Section~\ref{families} we explain how to create some of the most studied families of linear codes, including Hamming codes, cyclic codes, Reed-Solomon codes, and Reed-Muller codes. Finally, we give instructions on how to create locally recoverable codes.

In this paper we do not attempt to fully explain every function distributed in this package.
%It is important to mention that the functions are not fully explained in this paper. 
For a detailed explanation of all functions in the package, we refer to the Macaulay2 help page which can be accessed by
running 
%installPackage("CodingTheory", RemakeAllDocumentation=>true)
\begin{verbatim}
viewHelp CodingTheory
\end{verbatim}

@@

\section{Defining linear codes}\label{def}
Let $\mathbb{F}_q$ be a finite field with $q$ elements. Mathematically, a {\it linear code} is defined as a vector subspace $C\subseteq \mathbb{F}_q^n.$ For Macaulay2 (M2), a linear code is a submodule of $\mathbb{F}_q^n.$ Assume $q=p^r,$ where $p$ is a prime number and $r$ a positive integer. By definition, the dual code $C^\perp$ is the orthogonal complement of $C$ in $\mathbb{F}_q^n$ with respect to the standard inner product. One can define $C$ by specifying a list $L$ of elements of $\mathbb{F}_q^n$ that span $C$ or by giving a
a {\it generator matrix} $G$ whose rows form a basis of $C$. Alternatively, one can specify a list $L_{H}$ of elements of $\mathbb{F}_q^n$ that span the dual  code $C^\perp$ or a {\it parity check matrix} $H$ whose columns form a basis of the dual code $C^\perp$.
Below are the commands for the constructor linearCode to construct these equivalent instances of the LinearCode type:
\begin{itemize}
\item linearCode($\mathbb{F}_q$,$L$)
%\item linearCode($\mathbb{F}_q^n$, $L$)
%\item linearCode($\mathbb{F}_q^n$, $L_H$, ParityCheck $=>$ true)
%\item linearCode($\mathbb{F}_q$,$L_{H}$, ParityCheck $=>$ true)
\item linearCode($\mathbb{F}_q$,$n$,$L$)
\item linearCode($G$)
\item linearCode($\mathbb{F}_q$,$L_H$, ParityCheck $=>$ true)
\item linearCode($H$, ParityCheck $=>$ true)
\item linearCode($p$,$r$,$n$,$L$)
\item linearCode($p$,$r$,$n$,$L_H$, ParityCheck $=>$ true)
\end{itemize}

Now, here is a more specific example of how to construct a simple linear code:

\begin{example}
$\,$
\begin{verbatim}
i2 : F = GF 4;
i3 : L = {{1,1,0,0},{0,0,1,1}};
i4 : C = linearCode(F,L)
o4 = Code with Generator Matrix: | 1 1 0 0 |
                                 | 0 0 1 1 |
o4 : LinearCode
\end{verbatim}
\end{example}

One way to refer to a primitive element of a finite field is by specifying a symbol using the Variable option of the constructor GF.

\begin{example}
$\,$
\begin{verbatim}
i2 : F = GF(9,Variable => a);
i3 : LH = {{1,0,a,0,0},{0,a,a+1,1,0},{1,1,1,a,0}};
i4 : C = linearCode(F,LH,ParityCheck => true)
o4 = Code with Generator Matrix: | a-1 0 a+1 1 0 |
                                 | 0   0 0   0 1 |
o4 : LinearCode
\end{verbatim}
\end{example}
To construct a linear code from a matrix, it is necessary to correctly specify the underlying field. This can be done by passing a field to the matrix constructor.
\begin{example}
$\,$
\begin{verbatim}
i2 : F = GF 4;

i3 : M = matrix(F, {{1,0,1,0},{0,1,1,1}});

             2       4
o3 : Matrix F  <--- F

i4 : C = linearCode(M)

o4 = Code with Generator Matrix: | 1 0 1 0 |
                                 | 0 1 1 1 |

o4 : LinearCode

\end{verbatim}
\end{example}

\section{Basic parameters linear codes}~\label{pro}

The {\it dimension\/} and the {\it length\/} are two of the basic parameters of a code $C\subseteq \mathbb{F}_q^n$. They are defined as the subspace dimension $\dim_{\mathbb{F}_q} C$ and the ambient space dimension $n$, respectively. A third basic parameter is the {\it minimum weight}, which is given by 
\[
\min\{\|c\|
\colon  c \in C, c \neq 0\},
\]
where $\| c \|$ is the number of non-zero entries of $c$. The {\it rate} of $C$ is defined as the rational number $k/n.$ %The dual of $C$ is defined as the orthogonal complement of $C$ in $\mathbb{F}^n_q$ with respect to the usual dot product in $\mathbb{F}^n_q$.
 Some of the functions that can be used in M2 to compute basic parameters and algebraic objects associated with linear codes are the following: \newline
\begin{minipage}[t]{0.33\textwidth}
\begin{itemize}[leftmargin=1 cm]
\item C.GeneratorMatrix
\item C.Generators
\item C.ParityCheckMatrix
\item C.AmbientModule
\item alphabet C
\end{itemize}
\end{minipage}
\begin{minipage}[t]{0.33\textwidth}
\begin{itemize}
\item field C
\item informationRate C
\item ambientSpace C
\item length C
\item dim C
\end{itemize}
\end{minipage}
\begin{minipage}[t]{0.33\textwidth}
\begin{itemize}
\item minimumWeight C
\item codewords C
\item dualCode C
\item shorten(C, List)
\item ==
\end{itemize}
\end{minipage}

\begin{example}
$\,$
\begin{verbatim}
i2: F = GF 4;
i3 : L = {{1,1,0,0},{0,0,1,1}};
i4 : C = linearCode(F,L)
o4 = Code with Generator Matrix: | 1 1 0 0 |
                                 | 0 0 1 1 |
o4 : LinearCode
i5 : length C
o5 = 4
i6 : dim C
o6 = 2
i7 : informationRate C
     1
o7 = -
     2
o7 : QQ
i8 : ambientSpace C
      4
o8 = F
o8 : F-module, free
i9 : alphabet C
o9 = {0, a, a + 1, 1}
o9 : List
i10 : minimumWeight C
o10 = 2
i11 : dualCode C
o11 = Code with Generator Matrix: | 1 1 0 0 |
                                  | 0 0 1 1 |

o11 : LinearCode
\end{verbatim}
\end{example}

\section{Evaluation codes}\label{eva}
Let $\mathcal{X}=\left\{{\bf a}_1,\ldots,{\bf a}_n\right\}$ be a subset of an $m$-dimensional space $\mathbb{F}_q^{m}$.
Consider a finite dimensional subspace $S\subset {\mathbb{F}_q}[X_1,\ldots,X_m]$  of the $m$-variate polynomial ring over $\mathbb{F}_q$.
%and $S_{\leq d}$ the ${\mathbb{F}_q}$-vector space of all polynomials in $S$ of degree at most $d.$

The {\it evaluation map\/}
\begin{equation*}\label{functionevaluation}
{\rm ev}_S\colon S\longrightarrow {\mathbb{F}_q}^{|\mathcal{X}|},\ \ \ \ \ 
f\mapsto \left(f({\bf a}_1),\ldots,f({\bf a}_n)\right),
\end{equation*}
defines a linear map of ${\mathbb{F}_q}$-vector spaces. The image of ${\rm ev}_S$ in ${\mathbb{F}_q}^{|\mathcal{X}|}$, denoted by $C_{\mathcal{X}}(S)$, %defines a ${\mathbb{F}_q}$-vector subspace. We refer to $C_{\mathcal{X}}(d)$ 
is the {\it evaluation code\/} on the set $\mathcal{X}$ corresponding to $S$. The {\it vanishing ideal} of $\mathcal{X}$, denoted by $I(\mathcal{X})$, is the ideal in $S$ of all polynomials that vanish on  $\mathcal{X}.$ A key observation that allows the use of commutative algebra in studying evaluation codes is that the kernel of the evaluation map ${\rm ev}_S$ is precisely $S\cap I(\mathcal{X})$.

An evaluation code $C_{\mathcal{X}}(S)$ is defined in M2 as a separate type because there are more objects associated with it than with a linear code. For instance, the vanishing ideal associated to the set $\mathcal{X}$ plays an important role when finding and estimating parameters of the code, so it is convenient to be able to access it. Given an evaluation code $C$ in M2, the object C.linearCode is a linear code in M2. The command $\text{evaluationCode}(\mathbb{F}_q,\mathcal{X},L)$ defines an evaluation code where $\mathcal{X}$ is a list of elements in $\mathbb{F}_q^{m}$ and $L$ is a list of polynomials that span $S$. In the case when polynomials in $L$ are monomials, one may give the matrix of exponent vectors instead of $L$.

There are many construction of evaluation codes for specific choices of the set $\mathcal{X}$ and the subspace $S$. These include
Reed-Muller codes, Cartesian and monomial Cartesian codes, toric codes, and evaluation codes from graphs. We refer to 
\cite{carvalho4, Ha1, LSc, lopez-villa, MPV, algcodes, Ru, SoSo} for details on how these codes are defined and what properties they have from coding theory, commutative algebra, and algebraic geometry perspectives. Some functions defined in this package for various constructions of evaluation codes and associated algebraic objects are the following:

@@

\begin{minipage}[t]{0.52\textwidth}
\begin{itemize}[leftmargin=0.55 cm]
\item evaluationCode($\mathbb{F}_q$,List,List)
\item toricCode($\mathbb{F}_q$,Integer matrix)
\item cartesianCode($\mathbb{F}_q$,List,List)
\item orderCode($\mathbb{F}_q$,List,List,ZZ)
\item evCodeGraph($\mathbb{F}_q$,Incident Matrix,integer)
\end{itemize}
\end{minipage}
\begin{minipage}[t]{0.45\textwidth}
\begin{itemize}[leftmargin=1 cm]
\item vNumber(Ideal)
\item footPrint(Integer,Integer,Ideal)
\item hYpFunction(Integer,Integer,Ideal)
\item gMdFunction(Integer,Integer,Ideal)
\item vasFunction(Integer,Integer,Ideal)
\end{itemize}
\end{minipage}

@@

The mathematical definitions of the vNumber, the footprint function, the hyp function, the generalized footprint function and the Vasconcelos function can be found in~\cite{CSTVV}. The following example shows how to construct an evaluation code.

\begin{example}
$\,$
\begin{verbatim}
i2 : F=GF(4); R=F[x,y,z];
i4 : P={{0,0,0},{1,0,0},{0,1,0},{0,0,1},{1,1,1},{a,a,a}};
i5 : S={x+y+z,a+y*z^2,z^2,x+y+z+z^2};
i6 : C=evaluationCode(F,P,S)
o6 = Code with Generator Matrix: | 0 1 1 1 1   a   |
                                 | a a a a a+1 a+1 |
                                 | 0 0 0 1 1   a+1 |
                                 | 0 1 1 0 0   1   |
o6 : EvaluationCode
i7 : length C.LinearCode
o7 = 6
i8 : dim C.LinearCode
o8 = 3
i9 : C.Points
o9 = {{0, 0, 0}, {1, 0, 0}, {0, 1, 0}, {0, 0, 1}, {1, 1, 1}, {a, a, a}}
o9 : List
i10 : C.VanishingIdeal;
o10 = Ideal of R
\end{verbatim}
\end{example}
\section{Families of linear codes}\label{families}
We continue with the same notation: $n$ represents the length of the code, $k$ the dimension and $q$ the size of the field.
Some families of linear codes that have been implemented in this package are the following:

@@

\begin{minipage}[t]{0.5\textwidth}
\begin{itemize}[leftmargin=1 cm]
\item HammingCode($q$,integer)
\item randLDPC($n$, $k$, integer, integer)
\item cyclicCode($\mathbb{F}_q$ ,polynomial, $n$)
\item quasiCyclicCode($\mathbb{F}_q$,list)
\item RSCode($\mathbb{F}_q$,List,integer)
\item RMCode($q$,integer,integer)
\end{itemize}
\end{minipage}
\begin{minipage}[t]{0.5\textwidth}
\begin{itemize}[leftmargin=1 cm]
\item zeroCode($\mathbb{F}_q$,$n$)
\item universeCode($\mathbb{F}_q$,$n$)
\item repetitionCode($\mathbb{F}_q$,$n$)
\item zeroSumCode($\mathbb{F}_q$,$n$)
\item random($\mathbb{F}_q$, $n$, $k$)

\end{itemize}
\end{minipage}

@@

Mathematical definitions of the above families can be found in \cite{huf-pless,MacWilliams-Sloane,van-lint}.

>>>>

\begin{document}
\title{Coding theory package for {\it Macaulay2}}
%%%%%%%%%%%%%%%%%%%%%%%
\author{Taylor Ball}
\address[Taylor Ball]{Department of Mathematics\\University of Notre Dame\\ Notre Dame, IN USA}
\email{trball13@gmail.com}
%%%%%%%%%%%%%%%%%%%%%%%
\author{Eduardo Camps}
\address[Eduardo Camps]{Escuela Superior de F\'isica y Matem\'aticas \\ Instituto Polit\'ecnico Nacional\\ Mexico City, Mexico}
\email{camps@esfm.ipn.mx}
%%%%%%%%%%%%%%%%%%%%%%%
\author{Henry Chimal-Dzul}
\address[Henry Chimal-Dzul]{Department of Mathematics and Center of Ring Theory and its Applications\\ Ohio University\\ Athens, OH USA}
\email{hc118813@ohio.edu}
%%%%%%%%%%%%%%%%%%%%%%%
\author{Delio Jaramillo-Velez}
\address[Delio Jaramillo-Velez]{
Departamento de
Matem\'aticas\\
Centro de Investigaci\'on y de Estudios
Avanzados del
IPN\\
Apartado Postal
14--740 \\
07000 Mexico City, D.F.
}

@@

\begin{abstract}
In this {\it Macaulay2} \cite{M2} package we implement an object called {\tt LinearCode}. We implement functions that compute basic parameters and objects associated with a linear code, such as generator and parity check matrices, the dual code, length, dimension, and minimum distance, among others. We implement an object called {\tt EvaluationCode}, a construction which allows to study linear codes using tools of
algebraic geometry and commutative algebra. We implement functions to generate important families of linear codes such as Hamming codes, cyclic codes, Reed--Solomon codes, Reed--Muller codes, Cartesian codes, monomial--Cartesian codes, and toric codes. In addition, we implement functions for the syndrome decoding algorithm and locally recoverable code construction, which are important tools in applications of linear codes. The package \textit{CodingTheory} is available at \url{https://github.com/Macaulay2/Workshop-2020-Cleveland/tree/CodingTheory/CodingTheory}
\end{abstract}

@@

In this note we introduce the new {\it Macaulay2} \cite{M2} package called {\tt CodingTheory}. The goal of this package is to provide a range of functions for constructing  linear and evaluation codes over finite fields, and for computing some of their main properties. To this aim, we implement two types of objects, {\tt LinearCode} and {\tt EvaluationCode}. The package also includes implementation of functions for generating important families of linear codes like Hamming codes, cyclic codes, Reed-Solomon codes, Reed-Muller codes, Cartesian codes, monomial-Cartesian codes and toric codes. It also has functions for the syndrome decoding algorithm and locally recoverable codes.

The organization of this note is as follows. In Section~\ref{def} we describe various ways to construct a linear code over a finite field using the \textit{CodingTheory} package. In Section~\ref{pro} we show how to compute the main parameters of a linear code: length, dimension, and minimum distance. We also illustrate how to compute some of the main algebraic objects associated with linear codes like generator and parity check matrices, dual codes, etc. In Section~\ref{eva} we give a brief introduction to evaluation codes and describe some functions implemented to study these objects. In Section~\ref{families} we explain how to create some of the most studied families of linear codes, including Hamming codes, cyclic codes, Reed-Solomon codes, and Reed-Muller codes. Finally, we give instructions on how to create locally recoverable codes.

In this paper we do not attempt to fully explain every function distributed in this package. For a detailed explanation of all functions in the package, we refer to the {\it Macaulay2} help page which can be accessed by
running

\begin{verbatim}
i1: viewHelp CodingTheory
\end{verbatim}

@@

\section{Constructing linear codes}\label{def}

Let $\mathbb{F}_q$ be a finite field with $q$ elements. Mathematically, a {\it linear code} is defined as a vector subspace $C\subseteq \mathbb{F}_q^n$ and it is often specified by a {\it generator matrix}, which is a $k \times n$ matrix $G$ with entries in $\mathbb{F}_q$ whose $k$ rows form a basis for $C$. In {\it Macaulay2}, a linear code is defined as an $\mathbb{F}_q$-submodule of $\mathbb{F}_q^n$ using the constructor {\tt linearCode}.  This constructor is an instance of the {\tt LinearCode} type. There are various ways to use the command {\tt linearCode}. For example, one can use this command to construct a linear code $C\subseteq \mathbb{F}^n_q$ by specifying a generator matrix $G$ of $C$ (Example \ref{example_2.1}). Alternatively, one can use the command {\tt linearCode} to construct a linear code $C\subseteq \mathbb{F}_q^n$ by indicating the finite field $\mathbb{F}_q$ and a list $L$ of elements of $\mathbb{F}_q^n$ that span $C$ (Example \ref{example_2.2}).  More details and equivalent ways to use the constructor {\tt linearCode} are given next.

@@

\begin{minipage}[t]{0.6\textwidth}
 Inputs:
 \begin{itemize}[leftmargin=0.55 cm]
  \item {\tt F = GF(q)}, a finite field with {\tt q} elements
  \item {\tt n, r, p}, positive integers with {\tt p} prime
  \item {\tt G}, a matrix with entries in {\tt GF(q)}
  \item {\tt L}, a list of elements of $\mathtt{GF(q)}^\mathtt{n}$
 \end{itemize}
\end{minipage}
\begin{minipage}[t]{0.35\textwidth}
Usage:
\begin{itemize}[leftmargin=0.55 cm]
\item {\tt linearCode(G)}
\item {\tt linearCode(F,L)}
\item {\tt linearCode(F,n,L)}
\item {\tt linearCode(p,r,n,L)}
\end{itemize}
\end{minipage}

@@

In the next examples we construct a simple binary linear code $C\subseteq \mathbb{F}^4_2$ with generator matrix 
$G=
\begin{pmatrix}
 1 & 1 & 0 & 0\\
 0 & 0 & 1 & 1\\
\end{pmatrix}
$ using equivalent versions of the command {\tt linearCode}.

\begin{example}\label{example_2.1}
 $\,$
\begin{verbatim}
i1 : F = GF(2);
i2 : L = {{1,1,0,0},{0,0,1,1}};
i3 : G = matrix apply(L,codeword->apply(codeword,entry->sub(entry,F)));
i4 : C = linearCode(G);
i5 : C.GeneratorMatrix
o5 = | 1 1 0 0 |
     | 0 0 1 1 |
 \end{verbatim}
\end{example}
\vskip -0.5 cm
Note that in Example \ref{example_2.1} was necessary to coerce the entries of each vector in the list {\tt L} into elements of {\tt F = GF(2)}. An equivalent way to do this is to pass the field {\tt GF(q)} to the {\tt matrix} constructor or to use the constructor {\tt linearCode(GF(q),L)}.

\begin{example}\label{example_2.2}
 $\,$
\begin{verbatim}
i1 : L = {{1,1,0,0},{0,0,1,1}};
i2 : C = linearCode(GF(2),L);
i3 : C.GeneratorMatrix
o3 = | 1 1 0 0 |
     | 0 0 1 1 |
\end{verbatim}
\end{example}
The set $\mathbb{F}^{\ast}_q$ of non-zero elements of a finite field $\mathbb{F}_q$ is a multiplicative cyclic group (\cite[Theorem 3.3.1]{huf-pless}). A generator of $\mathbb{F}_q^{\ast}$ is called a \textit{primitive element} of $\mathbb{F}_q$. One way to refer to a primitive element of a finite field is by specifying a symbol using the {\tt Variable} option of the constructor {\tt GF}. In the next example we illustrate how to define a linear code  $C\subseteq \mathbb{F}_{11}^{10}$ with generator matrix
\[
 G=
 \begin{pmatrix}
  1 & a^1 & a^2 & \cdots & a^9\\
  1 & a^2 & a^4 & \cdots & (a^2)^9\\
  1 & a^3 & (a^3)^2 & \cdots & (a^3)^9\\
  1 & a^4 & (a^4)^2 & \cdots & (a^4)^9\\
 \end{pmatrix},
\]
where $a$ is a primitive element of $\mathbb{F}_{11}$. In {\it Macaulay2}, $a=2$.

\begin{example}\label{example_2.4}
 $\,$
\begin{verbatim}
i1 : F = GF(11,Variable => a);
i2 : G = matrix table({1,2,3,4},
                      {1,a,a^2,a^3,a^4,a^5,a^6,a^7,a^8,a^9},(i,j)->j^i);
i3 : C = linearCode(G);
i4 : C.GeneratorMatrix
o4 = | 1 2  4  -3 5  -1 -2 -4 3  -5 |
     | 1 4  5  -2 3  1  4  5  -2 3  |
     | 1 -3 -2 -5 4  -1 3  2  5  -4 |
     | 1 5  3  4  -2 1  5  3  4  -2 |
\end{verbatim}
\end{example}

In $\mathbb{F}_q^n$ there is a standard inner product defined for all $x=(x_1,\ldots,x_n)$ and $y=(y_1,\ldots,y_n)$ in $\mathbb{F}_q^n$ as $x\cdot y=x_1y_1+\cdots+x_ny_n.$ Given a linear code $C\subseteq \mathbb{F}^n_q$, the {\it dual} or {\it orthogonal code} of $C$ with respect to this product is the linear code $C^{\perp}$ defined as
\[
C^\perp=\{x\in \mathbb{F}_q^n\,:\,x\cdot c=0 \text{ for all }c\in C\}.
\]
A generator matrix of $C^{\perp}$ is called a {\it parity check matrix} for $C$. Since $(C^{\perp})^{\perp}=C$, another common way to mathematically define $C$ is by specifying one of its parity check matrices. By using the command {\tt linearCode} and the {\tt ParityCheck} option of this command, we can define a linear code $C\subseteq \mathbb{F}^n_q$ by either specifying a parity check matrix $H$ of $C$ or a list $L$ of elements of $\mathbb{F}_q^n$ that span $C^{\perp}$.

%We now present some examples to illustrate how to use the {\tt ParityCheck} option of the command {\tt linearCode}. 
In the next example, we define a linear code $C\subseteq \mathbb{F}_9^5$ whose dual code $C^{\perp}$ is generated by the set $\{(1,0,a,0,0),(0,1,a+1,1,0),(1,1,1,a,0)\}\subseteq \mathbb{F}^5_9$, where $a$ is a primitive element of $\mathbb{F}_9$. In {\it Macaulay2}, $a\in\mathbb{F}_9$ satisfies $a^2=a+1$.

\begin{example}\label{example_2.5}
$\,$
\begin{verbatim}
i1 : F = GF(9,Variable => a);
i2 : L = {{1,0,a,0,0},{0,a,a+1,1,0},{1,1,1,a,0}};
i3 : C = linearCode(F,L,ParityCheck => true);
i4 : C.GeneratorMatrix
o4 = | a-1 0 a+1 1 0 |
     | 0   0 0   0 1 |
i5 : C.ParityCheckMatrix
o5 = | 1 0 a   0 0 |
     | 0 a a+1 1 0 |
     | 1 1 1   a 0 |
\end{verbatim}
\end{example}

Although the dual code of a linear code can be constructed using the command {\tt dualCode} implemented in the {\it CodingTheory} package, the {\tt ParityCheck} option of the command {\tt linearCode} also allow us to construct the dual of a linear code. In the next example we construct the dual of the linear code in Example \ref{example_2.4}.

\begin{example}\label{example_2.5}
 $\,$
 \begin{verbatim}
i1 : F = GF(11,Variable => a);
i2 : G = matrix table({1,2,3,4},
                      {1,a,a^2,a^3,a^4,a^5,a^6,a^7,a^8,a^9},(i,j)->j^i);
i3 : D = linearCode(G,ParityCheck => true);
i4 : D.GeneratorMatrix
o4 = | 1  -3 5  3  1 0 0 0 0 0 |
     | -3 -1 4  -4 0 1 0 0 0 0 |
     | 4  -4 -3 5  0 0 1 0 0 0 |
     | -5 -3 4  4  0 0 0 1 0 0 |
     | -4 -4 -1 3  0 0 0 0 1 0 |
     | -3 5  3  1  0 0 0 0 0 1 |
 \end{verbatim}
\end{example}

\section{Basic parameters of linear codes}~\label{pro}

The {\it dimension\/}, {\it length\/}, and {\it rate} or {\it information rate}  of a linear code are basic parameters. Given a linear code $C\subseteq \mathbb{F}^n_q$, these parameters are defined as $k=\dim_{\mathbb{F}_q} C$, $n$, and $k/n$ respectively. Another parameter of a linear code $C\subseteq \mathbb{F}^n_q$ is the {\it minimum Hamming weight}, which is given by 
\[
w_H(C)=\min\{\|c\|
\colon  c \in C, c \neq 0\},
\]
where $\| c \|$ is the number of non-zero entries of $c\in \mathbb{F}^n_q$. This parameter is important in determining the error-correcting capability of $C$; the higher the minimum Hamming weight, the more errors the code can detect and correct (see \cite{huf-pless}). This function has been implemented \textcolor{red}{Nathan, write here something about the implementation of the weight function}.

In {\it Macaulay2}, the space $\mathbb{F}_q^n$ has been called the {\it ambient module} or {\it ambient space} of $C$, while the field $\mathbb{F}_q$, the {\it alphabet} or {\it field} of $C$. As often, the elements of $C$ are referred to as {\it codewords}. The input in all the following commands implemented in the {\tt CodingTheory} package is always a linear code  {\tt C}.

@@

\begin{minipage}[t]{0.3\textwidth}
\begin{itemize}[leftmargin=0.5 cm]
\item {\tt dim C}
\item {\tt length C}
\item {\tt alphabet C}
\item {\tt ambientSpace C}
\end{itemize}
\end{minipage}
\begin{minipage}[t]{0.32\textwidth}
\begin{itemize}[leftmargin=0.35 cm]
\item {\tt field C}
\item {\tt codewords C}
\item {\tt C.Generators}
\item {\tt C.GeneratorMatrix}
\end{itemize}
\end{minipage}
\begin{minipage}[t]{0.35\textwidth}
\begin{itemize}[leftmargin=0.55 cm]
\item {\tt C.AmbientModule}
\item {\tt minimumWeight C}
\item {\tt informationRate C}
\item {\tt C.ParityCheckMatrix}
\end{itemize}
\end{minipage}
\newpage
\begin{example}
$\,$
\begin{verbatim}
i1 : L = {{1,1,0,0},{0,0,1,1}};
i2 : C = linearCode(GF(4),L);
i3 : dim C
o3 : 2
i4 : length C
o4 = 4
i5 : alphabet C
o5 = {0, a, a + 1, 1}
i6 : ambientSpace C
           4
o6 = (GF 4)
i7 : field C
o7 = GF 4
i8 : minimumWeight C
o8 = 2
i9 : codewords C
o9 = {{1, 1, 1, 1}, {1, 1, a, a}, {a, a, 1, 1}, {a, a, a, a}, 
     ------------------------------------------------------------------
      {a + 1, a + 1, a, a}, {a + 1, a + 1, 1, 1}, {1, 1, a + 1, a + 1},
     ------------------------------------------------------------------
      {a, a, a + 1, a + 1}, {a + 1, a + 1, a + 1, a + 1}, {1, 1, 0, 0},
     ------------------------------------------------------------------
      {0, 0, 1, 1}, {0, 0, a, a}, {a, a, 0, 0}, {a + 1, a + 1, 0, 0}, 
     ------------------------------------------------------------------
      {0, 0, a + 1, a + 1}, {0, 0, 0, 0}}
\end{verbatim}
\end{example}
\section{Evaluation codes}\label{eva}
Let $P=\left\{{\bf a}_1,\ldots,{\bf a}_n\right\}$ be a subset of $\mathbb{F}_q^{m}$.
Consider a finite dimensional subspace $\mathcal{S}\subset {\mathbb{F}_q}[X_1,\ldots,X_m]$  of the ring of polynomials over $\mathbb{F}_q$ in $m$ variables. The {\it evaluation map\/}
\begin{equation*}\label{functionevaluation}
{\rm ev}_\mathcal{S}\colon \mathcal{S}\longrightarrow {\mathbb{F}_q}^{|P|},\ \ \ \ \ 
f\mapsto \left(f({\bf a}_1),\ldots,f({\bf a}_n)\right),
\end{equation*}
defines a linear map of ${\mathbb{F}_q}$-vector spaces. The image of ${\rm ev}_\mathcal{S}$ in ${\mathbb{F}_q}^{|P|}$, denoted by $C_{P}(\mathcal{S})$, %defines a ${\mathbb{F}_q}$-vector subspace. We refer to $C_{P}(d)$ 
is the {\it evaluation code\/} on the set $P$ corresponding to $\mathcal{S}$. The {\it vanishing ideal} of $P$, denoted by $I(P)$, is the ideal in $\mathbb{F}_q[X_1,\ldots,X_n]$ of all polynomials that vanish on  $P$. A key observation that allows the use of commutative algebra in studying evaluation codes is that the kernel of the evaluation map ${\rm ev}_S$ is precisely $\mathcal{S}\cap I(P)$.

An evaluation code $C_{P}(\mathcal{S})$ is implemented in {\it Macaulay2} as an {\tt EvaluationCode}. Although, the object {\tt C.LinearCode} is a linear code in {\it Macaulay2}. This has been done in this way because there are more objects associated with an evaluation code than with a linear code. For instance, the vanishing ideal associated to the set $P$ plays an important role when finding and estimating parameters of the code, so it is convenient to be able to access it.

There are many constructions of evaluation codes for specific choices of the set $P$ and the subspace $\mathcal{S}$. These include
Reed-Muller codes, Cartesian, monomial Cartesian codes, toric codes, and evaluation codes from graphs. We refer to 
\cite{carvalho4, Ha1, LSc, lopez-villa, MPV, algcodes, Ru, SoSo} for details on how these codes are defined and what properties they have from coding theory, commutative algebra, and algebraic geometry perspectives.

Some functions implemented in the {\it CodingTheory} package for various constructions of evaluation codes and associated algebraic objects are the following:

@@

Inputs:

\begin{minipage}[t]{0.53\textwidth}
\begin{itemize}[leftmargin=0.55 cm]
  \item {\tt I}, an ideal
  \item {\tt d,r}, positive integers
  \item {\tt F = GF(q)}, a finite field with {\tt q} elements
  \item {\tt P}, a list of points in $\mathtt{F}^\mathtt{m}$
  \item {\tt S}, a list of polynomials in {\tt m} variables
\end{itemize}
\end{minipage}
\begin{minipage}[t]{0.45\textwidth}
\begin{itemize}[leftmargin=0.55 cm]
  \item {\tt M}, an integer matrix
  \item {\tt L}, a list of subsets of {\tt F}
  \item {\tt v}, a list of {\tt m} positive integers
  \item {\tt MI}, an incident matrix of a graph
 \end{itemize}
\end{minipage}

Usage:

\begin{minipage}[t]{0.53\textwidth}
\begin{itemize}[leftmargin=0.55 cm]
\item {\tt evaluationCode(F,P,S)}
\item {\tt toricCode(F,M)}
\item {\tt cartesianCode(F,L,d)}
\item {\tt orderCode(F,P,v,d)}
\item {\tt evCodeGraph(F,MI,S)}
\end{itemize}
\end{minipage}
\begin{minipage}[t]{0.45\textwidth}
\begin{itemize}[leftmargin=0.55 cm]
\item {\tt vNumber(I)}
\item {\tt footPrint(d,r,I)}
\item {\tt hYp(d,r,I)}
\item {\tt genMinDisIdeal(d,r,I)}
\item {\tt vasconcelosDegree(d,r,I)}
\end{itemize}
\end{minipage}

The mathematical definitions of the functions in the second column of the previous list can be found in \cite{CSTVV}. The following example shows how to construct an evaluation code in {\it Macaulay2} using the {\it CodingTheory} package.

\begin{example}
$\,$
\begin{verbatim}
i1 : F=GF(4,Variable=>a); R=F[x,y,z];
i3 : P={{0,0,0},{1,0,0},{0,1,0},{0,0,1},{1,1,1},{a,a,a}};
i4 : S={x+y+z,a+y*z^2,z^2,x+y+z+z^2};
i5 : C=evaluationCode(F,P,S);
i6 : (C.LinearCode).GeneratorMatrix
o6 = | 0 1 1 1 1   a   |
     | a a a a a+1 a+1 |
     | 0 0 0 1 1   a+1 |
     | 0 1 1 0 0   1   |

i7 : length C.LinearCode
o7 = 6
i8 : dim C.LinearCode
o8 = 3
i9 : C.Points
o9 = {{0, 0, 0}, {1, 0, 0}, {0, 1, 0}, {0, 0, 1}, {1, 1, 1}, {a, a, a}}

i10 : C.VanishingIdeal

                         2    2                      2    2              
o10 = ideal (x*z + y*z, y  + z  + y + z, x*y + y*z, x  + z  + x + z, 
      --------------------------------------------------------------
      3           2
     z  + (a +1)z  + a*z)
\end{verbatim}
\end{example}
